---
title: 'Appendix S2: Landsat Filters'
subtitle: 'lakeCoSTR: An open-source, interactive retrieval tool to facilitate use of the Landsat Collection 2 surface temperature product to estimate lake surface water temperatures'
author: "*Herrick, C, Steele, BG*,  Brentrup, JA, Cook, B, Cottingham, KL, Ducey, M, Johnson, K, Lutz, DA, Palace, M, Sullivan, F, Thompson, M, Trout-Haney, JV, Weathers, KC"
date: 'Ecosphere'
output: word_document
editor_options: 
  markdown: 
    wrap: 90
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# read in libraries and functions
source('R_library.R')

#print r version
R.version$version.string

# point to directories
datadir = '~/GitHub/ids-ne-lakes/data/'
C2_datadir <- '~/GitHub/ids-ne-lakes/data/colab-output/C2/'
```

# Background:

Upon inspection of the lakeCoSTR-exported histograms, it was clear that
some of the ranges of data seemed ecologically unfeasible,
displaying a range of 20 degrees Celsius or more over the surface of Lake
Sunapee. We assume that these types of spread are indicative of lingering 
atmospheric correction issues or cloud effects. In this appendix, we explore some of the additional quality
assurance filters that we tried that may be helpful for others using
this tool.

![Landsat 4:
1990-07-21](C:/Users/steeleb/Documents/GitHub/ids-ne-lakes/data/colab-output/C2/histograms/LT04_L1TP_013030_19900721_20200915_02_T1_histo.png){width="3in"}![Landsat
5:
1998-07-23](C:/Users/steeleb/Documents/GitHub/ids-ne-lakes/data/colab-output/C2/histograms/LT05_L1TP_013030_19880723_20200917_02_T1_histo.png){width="3in"}

![Landsat 7:
2018-05-31](C:/Users/steeleb/Documents/GitHub/ids-ne-lakes/data/colab-output/C2/histograms/LE07_L1TP_013030_20180531_20200829_02_T1_histo.png){width="3in"}![Landsat
8:
2020-09-17](C:/Users/steeleb/Documents/GitHub/ids-ne-lakes/data/colab-output/C2/histograms/LC08_L1TP_013030_20200917_20201005_02_T1_histo.png){width="3in"}

Figure S1. Four histograms, one from each Landsat Mission, exhibiting unusual frequency distributions as exported from the lakeCoSTR tool. 

In addition to the possible range-of-value issues shown in Figure S1, 
there is also evidence of non-unimodal distributions. Because we were 
intersted in using a single median value to describe the lake surface 
temperature in a scene, we determined these scenes were not suitable for
our analysis.  Note that bimodal
distributions in surface temperature across a lake surface are certainly
ecologically possible outcomes and may be true for some systems or useful for some
analyses. Finally, There were also scenes
with negative temperatures reported even though pixels classified as
snow or ice were filtered out in the Colab script. 

Our desire was to create a filter that would remove the Landsat scenes with presumed atmospheric
interference (indicated by grossly large estimated temperature ranges) or those that were otherwise 
not suitable for our analysis. We used measurements from the extensive *in-situ* data network at 
Lake Sunapee to define some of these filters. We used statistical measures of distribution, including 
quartile values, spread, and measurement of distribution kurtosis (the
'tailedness' of the distribution). Knowing that we did not filter for
clouds or cloud shadows explicitly in the tool, and given that there is documented interference
with the surface temperature product (Cook, et al., 2014), we also tried a cloud filter to
eliminate some scenes.

# Methods

The filters we explored, listed in order of increased stringency
based on the number of scenes eliminated from analysis, were:

-   **freeze**: removing all scenes whose minimum temperature was below
    0 degrees Celsius

-   **IQR**: removing scenes that reported
    interquartile temperature ranges greater than 110% in a summary of the *in-situ* temperature
    record

-   **kurtosis**: removing any scenes whose histogram has a kurtosis value
    less than 2

-   **cloud**: removing scenes with cloud cover
    greater than 40%

-   **range**: removing scenes that reported temperature ranges
    greater than 110% observed in a summary of the *in-situ* temperature record

## Load, summarize, and filter data

```{r load data}
#read in previously collated and filtered in-situ data
insitu <- read.csv(paste0(datadir, 'insitu_temp_data_v2021-10-20.csv')) %>% 
  filter(!is.na(lat_dd)) %>% 
  mutate(datetime = as.POSIXct(datetime, tz = 'Etc/GMT+5'))

#read in Collection 2 paired data, an output from the Colab Notebook
C2ST <- read.csv(file.path(C2_datadir,paste0('C2_v2_temp_landsat_paired.csv')))
```

To define some filters, we used the validation dataset described in section 3.1.2 of the main text, filtered
to those temperatures measured between the hours of 9 and 11 am (the
approximate time of Landsat flyover). These values were aggregated to daily values of range, 
interquartile range, and number of locations contributing the ranges. From
these daily values, we calculated the maximum range and interquartile range from the
observed temperature data.

```{r}
# Summarize in-situ to define spread and IQR limits ####

# get range per day during hours of flyover
range_per_date <- insitu %>% 
  mutate(hour = as.numeric(format(datetime, '%H')),
         date = as.Date(datetime)) %>% 
  filter(hour >= 9 & hour < 11) %>% 
  group_by(date) %>% 
  summarize(temp_range = max(temp_degC) - min(temp_degC),
            IQR = IQR(temp_degC, na.rm = T),
            n_locs = length(unique(location)))

#save max observed spread and IQR for further filtering
max_spread <- max(range_per_date$temp_range, na.rm = T)
print('Maximium spread observed is:')
max_spread
max_IQR <- max(range_per_date$IQR, na.rm = T)
print('Maximium interquartile range observed is:')
max_IQR
```

```{r flag for spread and make datasets}
# Flag C2 data for freezing temps, spread P/F, IQR P/F ####
C2ST <- C2ST %>% 
  mutate(freeze_QAQC = case_when(surface_temp_min < 0 ~ 'F', #pass/fail for freezing temps
                                 TRUE ~ 'P')) %>% 
  mutate(temp_spread = round(surface_temp_max, digits = 1) - round(surface_temp_min, digits = 1)) %>% #calc temp spread; pass/fail for temp spread
  mutate(spread_QAQC = case_when(temp_spread < max_spread*1.1 ~ 'P',
                                 TRUE ~ 'F')) %>% 
  mutate(IQR = round(surface_temp_p75, digits = 1) - round(surface_temp_p25, digits = 1))  %>%  #calc IQR; pass/fail for temp IQR
  mutate(IQR_QAQC = case_when(IQR < max_IQR*1.1 ~ 'P',
                              TRUE ~ 'F'))

# filtered datasets
C2ST_freeze <- C2ST %>% 
  filter(freeze_QAQC == 'P')

C2ST_maxrange <- C2ST %>% 
  filter(spread_QAQC == 'P')

C2ST_maxIQR <- C2ST %>% 
  filter(IQR_QAQC == 'P')

C2ST_cloud <- C2ST %>% 
  filter(cloud_cover <40)

C2ST_kurtosis <- C2ST %>% 
  filter(surface_temp_kurtosis > 2)
```

```{r perform deming regresions}
# Deming Regressions ####

# deming regression for C2
C2_deming = deming::deming(C2ST$surface_temp_median ~ C2ST$is_temp_med)
C2_deming_forresid = mcreg(x = C2ST$is_temp_med, y = C2ST$surface_temp_median, method.reg = 'Deming')
C2ST$opt_resid = MCResult.getResiduals(C2_deming_forresid)$optimized
C2ST$filter = 'none'

# deming regression for C2 without freezing temps
C2_freeze_deming = deming::deming(C2ST_freeze$surface_temp_median ~ C2ST_freeze$is_temp_med)
C2_freeze_deming_forresid = mcreg(x = C2ST_freeze$is_temp_med, 
                           y = C2ST_freeze$surface_temp_median, 
                           method.reg = 'Deming')
C2ST_freeze$opt_resid = MCResult.getResiduals(C2_freeze_deming_forresid)$optimized
C2ST_freeze$filter = 'freeze'

# deming regression for C2 without freezing temps and within range
C2_maxrange_deming = deming::deming(C2ST_maxrange$surface_temp_median ~ C2ST_maxrange$is_temp_med)
C2_maxrange_deming_forresid = mcreg(x = C2ST_maxrange$is_temp_med, 
                                  y = C2ST_maxrange$surface_temp_median, 
                                  method.reg = 'Deming')
C2ST_maxrange$opt_resid = MCResult.getResiduals(C2_maxrange_deming_forresid)$optimized
C2ST_maxrange$filter = 'range'

# deming regression for C2 without freezing temps and within IQR
C2_maxIQR_deming = deming::deming(C2ST_maxIQR$surface_temp_median ~ C2ST_maxIQR$is_temp_med)
C2_maxIQR_deming_forresid = mcreg(x = C2ST_maxIQR$is_temp_med, 
                                  y = C2ST_maxIQR$surface_temp_median, 
                                  method.reg = 'Deming')
C2ST_maxIQR$opt_resid = MCResult.getResiduals(C2_maxIQR_deming_forresid)$optimized
C2ST_maxIQR$filter = 'IQR'

# deming regression for C2 without freezing temps and within IQR and data from the buoy only
C2_cloud_deming = deming::deming(C2ST_cloud$surface_temp_median ~ 
                                        C2ST_cloud$is_temp_med)
C2_cloud_deming_forresid = mcreg(x = C2ST_cloud$is_temp_med, 
                                  y = C2ST_cloud$surface_temp_median, 
                                  method.reg = 'Deming')
C2ST_cloud$opt_resid = MCResult.getResiduals(C2_cloud_deming_forresid)$optimized
C2ST_cloud$filter = 'cloud'

#deming regress for C2 with kurtosis >= 2
C2_kurtosis_deming = deming::deming(C2ST_kurtosis$surface_temp_median ~ 
                                   C2ST_kurtosis$is_temp_med)
C2_kurtosis_deming_forresid = mcreg(x = C2ST_kurtosis$is_temp_med, 
                                 y = C2ST_kurtosis$surface_temp_median, 
                                 method.reg = 'Deming')
C2ST_kurtosis$opt_resid = MCResult.getResiduals(C2_kurtosis_deming_forresid)$optimized
C2ST_kurtosis$filter = 'kurtosis'
```

# Presentation and discussion of filter performance

To analyze the performance of each of the filters, we performed a Deming regression (Deming, 1943) on each of the Landsat *in-situ* pair datasets filtered, as described above, from the output of the lakeCoSTR tool described in section 3.1.2 of the main text. 

The considerations we made when comparing filter performance were: 

1) number of valid scenes for validation
2) Pearson correlation coefficient
3) presence/absence of outliers
4) slope and intercept of the regression line

Scenes for validation ranged from 157 (no additional QAQC) to 71 (*in-situ* range filter). All datasets, including the unfiltered Collection 2 dataset, had acceptable Pearson correlation coefficients above 0.9 (Figure S2). The filters that reduced the outliers were the kurtosis filter, cloud filter, and the range filter. Of those, the kurtosis filter performed the best at reducing outliers. Of note, most outliers lie below the 1:1 line (Figure S2) - the presence of clouds and cloud shadows lowers the median temperature for each scene by reporting surface temperature values that are much cooler than in actuality.


```{r, fig.dim=c(10, 6)}
# Plot Deming regression for all filters to examine performance ####

DemingFigure_a <- ggplot(C2ST, aes(x = is_temp_med, y = surface_temp_median)) +
  geom_abline(slope = 1, intercept = 0, color = '#006cd1', size = 0.75) +
  geom_point() +
  #add deming regression and prediction intervals for C2
  geom_abline(intercept = C2_deming$coefficients[1], slope = C2_deming$coefficients[2], size = 0.75) +
  geom_abline(intercept = C2_deming$ci[1,1], slope = C2_deming$ci[2,1], linetype = 3, size = 0.75) +
  geom_abline(intercept = C2_deming$ci[1,2], slope = C2_deming$ci[2,2], linetype = 3, size = 0.75) +
  geom_text(label = paste0('r = ', round(cor(C2ST$surface_temp_median, C2ST$is_temp_med), digits = 3)),
            x = 2,
            y = 25,
            size = 4,
            hjust = 0)+
  geom_text(label = paste0('n = ', nrow(C2ST)),
            x = 2,
            y = 23,
            size = 4,
            hjust = 0)+
  labs(x = '',
       y = 'median Landsat-derived\nsurface temperature (deg C)',
       title = 'no filter') +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0, 27),
                  ylim = c(0, 27))

DemingFigure_b <- ggplot(C2ST_freeze, aes(x = is_temp_med, y = surface_temp_median)) +
  geom_abline(slope = 1, intercept = 0, color = '#006cd1', size = 0.75) +
  geom_point() +
  #add deming regression and prediction intervals for C2 filtered for freezing mins
  geom_abline(intercept = C2_freeze_deming$coefficients[1], slope = C2_freeze_deming$coefficients[2], size = 0.75) +
  geom_abline(intercept = C2_freeze_deming$ci[1,1], slope = C2_freeze_deming$ci[2,1], linetype = 3, size = 0.75) +
  geom_abline(intercept = C2_freeze_deming$ci[1,2], slope = C2_freeze_deming$ci[2,2], linetype = 3, size = 0.75) +
  geom_text(label = paste0('r = ', round(cor(C2ST_freeze$surface_temp_median, C2ST_freeze$is_temp_med), digits = 3)),
            x = 2,
            y = 25,
            size = 4,
            hjust = 0)+
  geom_text(label = paste0('n = ', nrow(C2ST_freeze)),
            x = 2,
            y = 23,
            size = 4,
            hjust = 0)+
  labs(x = '',
       y = '\n',
       title = 'freeze filter') +
    theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0, 27),
                  ylim = c(0, 27))

DemingFigure_c <- ggplot(C2ST_maxIQR, aes(x = is_temp_med, y = surface_temp_median)) +
  geom_abline(slope = 1, intercept = 0, color = '#006cd1', size = 0.75) +
  geom_point() +
  #add deming regression and prediction intervals for C2 filtered for sub zero and maxIQR
  geom_abline(intercept = C2_maxIQR_deming$coefficients[1], slope = C2_maxIQR_deming$coefficients[2], size = 0.75) +
  geom_abline(intercept = C2_maxIQR_deming$ci[1,1], slope = C2_maxIQR_deming$ci[2,1], linetype = 3, size = 0.75) +
  geom_abline(intercept = C2_maxIQR_deming$ci[1,2], slope = C2_maxIQR_deming$ci[2,2], linetype = 3, size = 0.75) +
  geom_text(label = paste0('r = ', round(cor(C2ST_maxIQR$surface_temp_median, C2ST_maxIQR$is_temp_med), digits = 3)),
            x = 2,
            y = 25,
            size = 4,
            hjust = 0)+
  geom_text(label = paste0('n = ', nrow(C2ST_maxIQR)),
            x = 2,
            y = 23,
            size = 4,
            hjust = 0)+
  labs(x = '',
       y = '\n',
       title = 'IQR filter') +
    theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0, 27),
                  ylim = c(0, 27))

DemingFigure_d <- ggplot(C2ST_kurtosis, aes(x = is_temp_med, y = surface_temp_median)) +
  geom_abline(slope = 1, intercept = 0, color = '#006cd1', size = 0.75) +
  geom_point() +
  #add deming regression and prediction intervals for C2 filtered for sub zero and kurtosis
  geom_abline(intercept = C2_kurtosis_deming$coefficients[1], slope = C2_kurtosis_deming$coefficients[2], size = 0.75) +
  geom_abline(intercept = C2_kurtosis_deming$ci[1,1], slope = C2_kurtosis_deming$ci[2,1], linetype = 3, size = 0.75) +
  geom_abline(intercept = C2_kurtosis_deming$ci[1,2], slope = C2_kurtosis_deming$ci[2,2], linetype = 3, size = 0.75) +
  geom_text(label = paste0('r = ', round(cor(C2ST_kurtosis$surface_temp_median, C2ST_kurtosis$is_temp_med), digits = 3)),
            x = 2,
            y = 25,
            size = 4,
            hjust = 0)+
  geom_text(label = paste0('n = ', nrow(C2ST_kurtosis)),
            x = 2,
            y = 23,
            size = 4,
            hjust = 0)+
  labs(x = '',
       y = 'median Landsat-derived\nsurface temperature (deg C)',
       title = 'kurtosis filter') +
    theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0, 27),
                  ylim = c(0, 27))

DemingFigure_e <- ggplot(C2ST_cloud, aes(x = is_temp_med, y = surface_temp_median)) +
  geom_abline(slope = 1, intercept = 0, color = '#006cd1', size = 0.75) +
  geom_point() +
  #add deming regression and prediction intervals for C2 filtered for sub zero and cloud
  geom_abline(intercept = C2_cloud_deming$coefficients[1], slope = C2_cloud_deming$coefficients[2], size = 0.75) +
  geom_abline(intercept = C2_cloud_deming$ci[1,1], slope = C2_cloud_deming$ci[2,1], linetype = 3, size = 0.75) +
  geom_abline(intercept = C2_cloud_deming$ci[1,2], slope = C2_cloud_deming$ci[2,2], linetype = 3, size = 0.75) +
  geom_text(label = paste0('r = ', round(cor(C2ST_cloud$surface_temp_median, C2ST_cloud$is_temp_med), digits = 3)),
            x = 2,
            y = 25,
            size = 4,
            hjust = 0)+
  geom_text(label = paste0('n = ', nrow(C2ST_cloud)),
            x = 2,
            y = 23,
            size = 4,
            hjust = 0)+
  labs(x = expression(bold(paste(italic('in-situ'), ' median water temp (deg C)'))),
       y = '\n',
       title = 'cloud filter') +
    theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0, 27),
                  ylim = c(0, 27))

DemingFigure_f <- ggplot(C2ST_maxrange, aes(x = is_temp_med, y = surface_temp_median)) +
  geom_abline(slope = 1, intercept = 0, color = '#006cd1', size = 0.75) +
  geom_point() +
  #add deming regression and prediction intervals for C2 filtered for sub zero and maxrange
  geom_abline(intercept = C2_maxrange_deming$coefficients[1], slope = C2_maxrange_deming$coefficients[2], size = 0.75) +
  geom_abline(intercept = C2_maxrange_deming$ci[1,1], slope = C2_maxrange_deming$ci[2,1], linetype = 3, size = 0.75) +
  geom_abline(intercept = C2_maxrange_deming$ci[1,2], slope = C2_maxrange_deming$ci[2,2], linetype = 3, size = 0.75) +
  geom_text(label = paste0('r = ', round(cor(C2ST_maxrange$surface_temp_median, C2ST_maxrange$is_temp_med), digits = 3)),
            x = 2,
            y = 25,
            size = 4,
            hjust = 0)+
  geom_text(label = paste0('n = ', nrow(C2ST_maxrange)),
            x = 2,
            y = 23,
            size = 4,
            hjust = 0)+
  labs(x = '',
       y = '\n',
       title = 'range filter') +
    theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0, 27),
                  ylim = c(0, 27))


DemingFigure <- plot_grid(DemingFigure_a, DemingFigure_b, DemingFigure_c, DemingFigure_d, DemingFigure_e, DemingFigure_f,
                   ncol = 3,
                   labels = c('a', 'b', 'c', 'd', 'e', 'f', '', 'g', ''),
                   label_x = 0.15)

DemingFigure
```

Figure S2. Deming regressions for 6 datasets of Landsat Collection 2 surface temperature product, arranged as least stringent to most stringent based on the number of scenes eliminated from the validation dataset. The blue line is the 1:1 line, the black line shows the Deming error-in-variables regression, and the dotted black lines indicate the 95% prediction intervals for the Deming regression. The Pearson correlation (r) and sample size (n) in the upper left corner.

While increasing the stringency of Landsat scene filters generally led to increased agreement between the Landsat-derived surface temperature product and the in-situ water temperature (Figure S3), it is notable that the filter that eliminated the most scenes from the analysis (“range”) has an estimated slope and intercept furthest from 1 and 0, respectively. Here, the kurtosis filter and the cloud filter performance was very similar. This may indicate that both filters perform similar functions.

```{r, fig.dim = c(8, 4)}
# Plot slope and intercept with 95%ci ####
slope_int_table <- NULL

slope_int_table$model = c( 'C2ST', 'C2ST_freeze', 'C2ST_maxrange', 'C2ST_kurtosis', 'C2ST_maxIQR', 'C2ST_cloud')
slope_int_table$slope = c(C2_deming$coefficients[2],
                          C2_freeze_deming$coefficients[2],
                          C2_maxrange_deming$coefficients[2],
                          C2_kurtosis_deming$coefficients[2],
                          C2_maxIQR_deming$coefficients[2],
                          C2_cloud_deming$coefficients[2])
slope_int_table$intercept = c(C2_deming$coefficients[1],
                              C2_freeze_deming$coefficients[1],
                              C2_maxrange_deming$coefficients[1],
                              C2_kurtosis_deming$coefficients[1],
                              C2_maxIQR_deming$coefficients[1],
                              C2_cloud_deming$coefficients[1])
slope_int_table$slope_se = c(C2_deming$se[2],
                             C2_freeze_deming$se[2],
                             C2_maxrange_deming$se[2],
                             C2_kurtosis_deming$se[2],
                             C2_maxIQR_deming$se[2],
                             C2_cloud_deming$se[2])
slope_int_table$int_se = c(  C2_deming$se[1],
                             C2_freeze_deming$se[1],
                             C2_maxrange_deming$se[1],
                             C2_kurtosis_deming$se[1],
                             C2_maxIQR_deming$se[1],
                             C2_cloud_deming$se[1])
slope_int_table <- as.data.frame(slope_int_table)

slope_int_table <- slope_int_table %>% 
  pivot_longer(cols = c(slope, intercept), names_to = 'var', values_to = 'value') %>% 
  mutate(se = case_when(var == 'slope' ~ slope_se,
                        var == 'intercept' ~ int_se,
                        TRUE ~ NA_real_),
         u95 = value + se,
         l95 = value - se) %>% 
  select(-slope_se, -int_se) %>% 
  mutate(model = factor(model, 
                        levels = c('C2ST', 'C2ST_freeze','C2ST_maxIQR', 'C2ST_kurtosis', 'C2ST_cloud', 'C2ST_maxrange'),
                        labels = c('no filter', 'freeze filter', 'IQR filter', 'kurtosis filter', 'cloud filter','range filter'))) %>% 
  mutate(regression = 'Deming')

slope_fig <- slope_int_table %>% 
  filter(var == 'slope') %>% 
  ggplot(., aes(x = model, y = value)) +
  geom_point() +
  geom_pointrange(aes(ymin = l95, ymax = u95))+
  geom_abline(intercept = 1, slope = 0, color = '#454545', lty=2) +
  labs(x = NULL,
       y = 'estimated slope') +
  coord_cartesian(ylim = c(0.85, 1.15)) +
  theme_bw()
int_fig <- slope_int_table %>% 
  filter(var == 'intercept') %>% 
  ggplot(., aes(x = model, y = value)) +
  geom_point() +
  geom_pointrange(aes(ymin = l95, ymax = u95))+
  geom_abline(intercept = 0, slope = 0, color = '#454545', lty=2) +
  labs(x = NULL,
       y = 'estmiated intercept') +
  coord_cartesian(ylim = c(-3.2, 0)) +
  theme_bw()

slope_int_fig <- plot_grid(slope_fig, int_fig,
                           ncol = 1,
                           labels = c('a', 'b'),
                           label_size = 12)
slope_int_fig
```
Figure S3. A simple comparison of estimated slope (a) and intercept (b) with upper and lower 95% confidence intervals for each of the Collection 2 (raw + filtered) datasets presented in Figure C indicates that the cloud filter dataset’s Deming regression provided a slope closest to 1 and an intercept closest to 0 (indicated by the dashed line on each panel). Datasets are ordered by increasing stringency, as measured by the number of scenes in the validation dataset, from left to right. 


The two filters that seemed the most promising were the kurtosis and cloud filters. We carried out all analyses described in the main text with both filters. We chose the kurtosis filter after this excercise, because it removed a number of outliers in the long-term monthly temperature analysis that the cloud filter did not. We suspect that kurtosis, a measure of tailedness, was a successful filter because it incorporates shape and range in a single value, and we suspect these are the symptoms of cloud cover for this system. Additionally, the kurtosis filter can be applied to the lakeCoSTR dataset without *in-situ* data for validation. The kurtosis filter value (we used a value of 2), may have to be changed for other systems - analyses at other lakes that have *in-situ* data for validation will be needed to determine if the value we chose is transferable. 


# Literature Citations

Cook M, Schott JR, Mandel J, Raqueno N. Development of an Operational Calibration Methodology for the Landsat Thermal Data Archive and Initial Testing of the Atmospheric Compensation Component of a Land Surface Temperature (LST) Product from the Archive. Remote Sensing. 2014; 6(11):11244-11266. https://doi.org/10.3390/rs61111244

Deming, W. E. 1943. Statistical adjustment of data. Dover Publications, 1985.