---
title: "Sunapee Temp and Landsat LST Figures and Tables"
author: "B. Steele"
date: "24Sept2021"
output: html_document
---

This script walks through the analyses and figures presented in the Herrick, et al, MS using all historical LS data for Lake Sunapee. 



```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

source('R_library.R')

dir = '~/GitHub/ids-ne-lakes/'
datadir = '~/GitHub/ids-ne-lakes/data/'
figdir = '~/GitHub/ids-ne-lakes/figures/'

```


#Read in all data

```{r}
#load in all high-frequency insitu data for historical data analysis
insitu <- read.csv(paste0(datadir, 'insitu_temp_data_v2021-05-17.csv')) %>% 
  filter(!is.na(lat_dd)) %>% 
  mutate(datetime = as.POSIXct(datetime, tz='Etc/GMT+5')) %>% 
  filter(as.numeric(format(datetime, '%m')) < 11 & as.numeric(format(datetime, '%m')) > 5)

#read in all skin temp data for whole lake
lst_all <- read_csv(paste0(datadir, 'colab-output/2021-06-08/sunapee_temp_stats.csv'),
                             col_types = c('')) %>% 
  mutate(date =as.Date(substrRight(`system:index`, 8), '%Y%m%d'),
         month = as.numeric(format(date, '%m')),
         year = as.factor(format(date, '%Y')),
         doy = as.numeric(format(date, '%j')),
         LSmission = case_when(grepl('LT04', `system:index`) ~ 'LS 4',
                               grepl('LT05', `system:index`) ~ 'LS 5',
                               grepl('LE07', `system:index`) ~ 'LS 7',
                               grepl('LC08', `system:index`) ~ 'LS 8')) %>% 
  filter(month < 11 & month > 5) %>%
  rowid_to_column()

lst_all %>% 
  group_by(LSmission) %>% 
  summarize(count = length(LSmission))

lst_paired <- read_csv(paste0(datadir, 'colab-output/2021-06-08/sunapee_insitu_landsat_paired.csv'),
                             col_types = c('')) %>% 
  mutate(date =as.Date(substrRight(`system:index`, 8), '%Y%m%d')) %>% 
  filter(as.numeric(format(date, '%m')) < 11 & as.numeric(format(date, '%m')) >5)

#load in the matched high-frequency insitu data from the .csv's outputted in the Google Colab script

#make a list of the file names in the folder
match_names <- list.files(paste0(datadir, 'colab-output/2021-06-08/ancillary'))

#iterate over the files into a dataframe
for(i in 1:length(match_names)) {
  data <- read_csv(paste0(datadir, 'match-files/', match_names[i]), 
           col_types = c('')) %>% 
    mutate(source = paste0(match_names[i]))
  if (i == 1) {
    match_details <- data
  } else {
  match_details <- full_join(match_details, data)
  }
}

# add columns for match_details
match_details <- match_details %>% 
  mutate(date =as.Date(datetime),
         month = as.numeric(format(date, '%m')),
         year = as.factor(format(date, '%Y')),
         doy = as.numeric(format(date, '%j'))) %>% 
  filter(frequency == 'sub-daily') %>% 
  filter(month < 11 & month > 5) 

#read in the modal information
modal_analysis <- read.csv(file.path(datadir, 'colab-output/2021-06-08/date_modalanalysis.csv')) %>% 
  mutate(date = as.Date(date),
         month =  as.numeric(format(date, '%m'))) %>% 
  filter(month < 11 & month > 5)

#read in binned data
lst_hist <- read.csv(file.path(datadir, 'colab-output/2021-06-08/sunapee_histograms_0p1bin_reformatted.csv')) %>% 
  mutate(date = as.Date(date),
         month =  as.numeric(format(date, '%m'))) %>% 
  filter(month < 11 & month > 5)

```


## Summary Data Tables for insitu database

```{r}

#number of observations for total DB
insitu %>% 
  summarize(nobs = length(datetime),
            ndays = length(unique(format(datetime, '%Y-%m-%d'))),
            year_start = min(as.numeric(format(datetime, '%Y'))),
            year_end = max(as.numeric(format(datetime, '%Y'))), 
            min_month = min(as.numeric(format(datetime, '%m'))),
            max_month = max(as.numeric(format(datetime, '%m'))),
            nsites = length(unique(location)))

#number of observations between time of 9 and 11
insitu %>% 
  mutate(hour = as.numeric(format(datetime, '%H'))) %>% 
  filter(hour >= 9 & hour < 11) %>% 
  summarize(nobs = length(datetime),
            ndays = length(unique(format(datetime, '%Y-%m-%d'))),
            year_start = min(as.numeric(format(datetime, '%Y'))),
            year_end = max(as.numeric(format(datetime, '%Y'))), 
            min_month = min(as.numeric(format(datetime, '%m'))),
            max_month = max(as.numeric(format(datetime, '%m'))),
            nsites = length(unique(location)))


#number of sites
length(unique(insitu$location))
unique(insitu$location)

#document max spread and IQR for LST filtering
range_per_date <- insitu %>% 
  mutate(hour = as.numeric(format(datetime, '%H')),
         date = as.Date(datetime)) %>% 
  filter(hour >= 9 & hour < 11) %>% 
  group_by(date) %>% 
  summarize(temp_range = max(temp_degC) - min(temp_degC),
            IQR = IQR(temp_degC, na.rm = T),
            n_locs = length(unique(location)))
max_spread <- max(range_per_date$temp_range, na.rm = T)
max_IQR <- max(range_per_date$IQR, na.rm = T)

range_summary <- range_per_date %>% 
  mutate(month = as.numeric(format(date, '%m'))) %>% 
  group_by(month) %>% 
  filter(n_locs >3) %>% 
  summarise(min_range = min(temp_range, na.rm = T),
            max_range = max(temp_range, na.rm = T))

```

## Summary of location/depth/measurement

```{r}
insitu %>% 
  mutate(year = format(datetime, '%Y'),
         month = as.numeric(format(datetime, '%m'))) %>% 
  group_by(year, location) %>% 
  summarise(min_depth_m = min(depth_m),
            max_depth_m = max(depth_m),
            min_month = min(month),
            max_month = max(month)) %>% 
  pivot_longer(!c(year, location), names_to = 'variable', values_to = 'value') %>% 
  pivot_wider(id_cols = c(year, variable, value),
              names_from = c(location, variable)) %>% 
  write.csv(., file.path(datadir, 'insitu_data_extent_summary.csv'), row.names = F)

```
 
## QAQC flags for freezing temp images; non unimodal; max spread or IQR <110% observed; Summary of lst data according to QAQC flags

```{r}
lst_all_val <- lst_all %>% 
  mutate(freeze_QAQC = case_when(surface_temp_min < 0 ~ 'F', #pass/fail for freezing temps
                                 TRUE ~ 'P')) %>% 
  left_join(., modal_analysis) %>% #join with modal analysis
  mutate(unimodal_QAQC = case_when(unimodal == 'unimodal' ~ 'P',
                                   TRUE ~ 'F')) %>%  # pass/fail for unimodality
  mutate(temp_spread = round(surface_temp_max, digits = 1) - round(surface_temp_min, digits = 1)) %>% #calc temp spread; pass/fail for temp spread
  mutate(spread_QAQC = case_when(temp_spread < max_spread*1.1 ~ 'P',
                                 TRUE ~ 'F')) %>% 
  mutate(IQR = round(surface_temp_p75, digits = 1) - round(surface_temp_p25, digits = 1))  %>%  #calc IQR; pass/fail for temp IQR
  mutate(IQR_QAQC = case_when(IQR < max_IQR*1.1 ~ 'P',
                              TRUE ~ 'F'))

lst_filt_maxrange <- lst_all_val %>% 
  filter(freeze_QAQC == 'P' & unimodal_QAQC == 'P' & spread_QAQC == 'P')

lst_filt_maxrange %>% 
  group_by(LSmission) %>% 
  summarize(count = length(LSmission))

lst_filt_maxIQR <- lst_all_val %>% 
  filter(freeze_QAQC == 'P' & unimodal_QAQC == 'P' & IQR_QAQC == 'P')

lst_filt_maxIQR %>% 
  group_by(LSmission) %>% 
  summarize(count = length(LSmission))

ggplot(lst_all_val, aes(x = pct_lake, y = (surface_temp_max-surface_temp_min))) +
  geom_point(aes(color = month)) +
  labs(y = 'temp_spread')+
  geom_abline(intercept = 1.1 * max_spread, slope = 0, lty = 2) +
  final_theme
ggsave(file.path(figdir, 'tempspread by pct lake by month summary.png'))

ggplot(lst_all_val, aes(x = pct_lake, y = (surface_temp_p75-surface_temp_p25))) +
  geom_point(aes(color = month)) +
  labs(y = 'temp_IQR')+
    geom_abline(intercept = 1.1 * max_IQR, slope = 0, lty = 2) +
final_theme
ggsave(file.path(figdir, 'temp IQR by pct lake by month summary.png'))

ggplot(lst_filt_maxrange, aes(x = pct_lake, y = temp_spread)) +
  geom_point(aes(color = month)) +
  final_theme
ggsave(file.path(figdir, 'filtered tempspread by pct lake by month summary.png'))

ggplot(lst_filt_maxIQR, aes(x = pct_lake, y = IQR)) +
  geom_point(aes(color = month)) +
  final_theme
ggsave(file.path(figdir, 'filtered temp IQR by pct lake by month summary.png'))

```

## Summary Data Tables for insitu-skin temp match

# LEFT OFF HERE TO MOVE FORWARD
```{r}
#add match details to lst_all_val
lst_all_val <- full_join(lst_all_val)

#inner join with unimodal data and 25% of lake
match_details_range <- match_details %>% 
  mutate(date = as.Date(date)) %>% 
  inner_join(., lst_all_val) %>% 
  filter(freeze_QAQC == 'P' & unimodal_QAQC == 'P' & spread_QAQC == 'P') %>% 
  filter(perc_lake > 25)

match_details_IQR <- match_details %>% 
  mutate(date = as.Date(date)) %>% 
  inner_join(., lst_all_val) %>% 
  filter(freeze_QAQC == 'P' & unimodal_QAQC == 'P' & IQR_QAQC == 'P') %>% 
  filter(perc_lake > 25)

#number of observations where there is a match with LS data
match_details_range %>% 
  summarize(nobs = length(datetime),
            ndays = length(unique(format(datetime, '%Y-%m-%d'))),
            year_start = min(as.numeric(format(datetime, '%Y'))),
            year_end = max(as.numeric(format(datetime, '%Y'))))

length(unique(match_details_range$location))

match_details_IQR %>% 
  summarize(nobs = length(datetime),
            ndays = length(unique(format(datetime, '%Y-%m-%d'))),
            year_start = min(as.numeric(format(datetime, '%Y'))),
            year_end = max(as.numeric(format(datetime, '%Y'))))

length(unique(match_details_IQR$location))

```

# Initial validation - using all data, by range and IQR

## pair qaqc skin temp with insitu as validation data
```{r}
lst_filt_mod_select_range <- lst_filt_maxrange %>% 
  select(rowid, date:temp_spread)
lst_filt_mod_select_IQR <- lst_filt_maxIQR %>% 
  select(rowid, date:IQR)

validation_range <- lst_paired %>% 
  mutate(date = as.Date(substrRight(scene, 8), '%Y%m%d')) %>% 
  inner_join(., lst_filt_mod_select_range)

validation_IQR <- lst_paired %>% 
  mutate(date = as.Date(substrRight(scene, 8), '%Y%m%d')) %>% 
  inner_join(., lst_filt_mod_select_IQR)

```

## predicted v observed by max range
```{r}
val_med_ls <- lm(validation_range$surface_temp_median ~ validation_range$temp_med)
summary(val_med_ls)

val_med_ls <-deming::deming(validation_range$surface_temp_median ~ validation_range$temp_med)
val_med_ls

cor(validation_range$surface_temp_median, validation_range$temp_med)

ggplot(validation_range, aes(x = temp_med, y = surface_temp_median)) +
  geom_point() +
  geom_abline(intercept = val_med_ls$coefficients[1], slope = val_med_ls$coefficients[2], color = 'blue') +
  geom_abline(intercept = val_med_ls$ci[1,1], slope = val_med_ls$ci[2,1], color = 'blue', linetype = 2) +
  geom_abline(intercept = val_med_ls$ci[1,2], slope = val_med_ls$ci[2,2], color = 'blue', linetype = 2) +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(xlim = c(7,27),
                  ylim = c(7,27)) +
  labs(y = 'satellite-derived median\nskin temperature (deg C)', 
       x = expression(paste(italic('in-situ'), ' median water temp (deg C)'))) +
  final_theme
ggsave(paste0(figdir, 'deming_range_v27Sept2021.jpg'), height = 5, width = 5, units = 'in')

validation_range$pred_range = val_med_ls$coefficients[1] + val_med_ls$coefficients[2]*validation_range$temp_med

lst_insitu_dem_forresid = mcreg(x = validation_range$temp_med, y = validation_range$surface_temp_median, method.reg = 'Deming')
validation_range$opt_resid_range = MCResult.getResiduals(lst_insitu_dem_forresid)$optimized

#plot to make sure prediction is correct
ggplot(validation_range, aes(x = temp_med, y = pred_range))+
  geom_point(color = 'blue') +
  geom_point(aes(x = temp_med, y = surface_temp_median))
```

### apply calibration of temp to LS-derived data
```{r}
validation_range <- validation_range %>% 
  mutate(calib_range = (surface_temp_median-val_med_ls$coefficients[1])/val_med_ls$coefficients[2]) %>% 
  select(rowid, temp_med, pred_range, opt_resid_range, calib_range)

lst_all_val <- lst_all_val %>% 
  full_join(., validation_range)

ggplot(lst_all_val, aes(x = temp_med, y = surface_temp_median)) +
  geom_point() +
  geom_point(aes(x = temp_med, y = calib_range), color = 'red')+
  geom_abline(intercept = val_med_ls$coefficients[1], slope = val_med_ls$coefficients[2], color = 'blue') +
  geom_abline(intercept = val_med_ls$ci[1,1], slope = val_med_ls$ci[2,1], color = 'blue', linetype = 2) +
  geom_abline(intercept = val_med_ls$ci[1,2], slope = val_med_ls$ci[2,2], color = 'blue', linetype = 2) +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(xlim = c(7,27),
                  ylim = c(7,27)) +
  labs(y = 'satellite-derived median\nskin temperature (deg C)', 
       x = expression(paste(italic('in-situ'), ' median water temp (deg C)'))) +
  final_theme


```
  

## predicted v observed by IQR
```{r}
val_med_IQR <- lm(validation_IQR$surface_temp_median ~ validation_IQR$temp_med)
summary(val_med_IQR)

val_med_IQR_dem <-deming::deming(validation_IQR$surface_temp_median ~ validation_IQR$temp_med)
val_med_IQR_dem

cor(validation_IQR$surface_temp_median, validation_IQR$temp_med)

ggplot(validation_IQR, aes(x = temp_med, y = surface_temp_median)) +
  geom_point() +
  geom_abline(intercept = val_med_IQR_dem$coefficients[1], slope = val_med_IQR_dem$coefficients[2], color = 'blue') +
  geom_abline(intercept = val_med_IQR_dem$ci[1,1], slope = val_med_IQR_dem$ci[2,1], color = 'blue', linetype = 2) +
  geom_abline(intercept = val_med_IQR_dem$ci[1,2], slope = val_med_IQR_dem$ci[2,2], color = 'blue', linetype = 2) +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(xlim = c(7,27),
                  ylim = c(7,27)) +
  labs(y = 'satellite-derived median\nskin temperature (deg C)', 
       x = expression(paste(italic('in-situ'), ' median water temp (deg C)'))) +
  final_theme
ggsave(paste0(figdir, 'deming_reg_median_IQR_v23Sept2021.jpg'), height = 5, width = 5, units = 'in')

validation_IQR$pred_IQR = val_med_IQR_dem$coefficients[1] + val_med_IQR_dem$coefficients[2]*validation_IQR$temp_med

lst_insitu_IQR_dem_forresid = mcreg(x = validation_IQR$temp_med, y = validation_IQR$surface_temp_median, method.reg = 'Deming')
validation_IQR$opt_resid_IQR = MCResult.getResiduals(lst_insitu_IQR_dem_forresid)$optimized

#plot to make sure prediction is correct
ggplot(validation_IQR, aes(x = temp_med, y = pred_IQR))+
  geom_point(color = 'blue') +
  geom_point(aes(x = temp_med, y = surface_temp_median))
```


### apply calibration of temp to LS-derived data - IQR
```{r}

validation_IQR<- validation_IQR %>% 
  mutate(calib_IQR = (surface_temp_median-val_med_IQR$coefficients[1])/val_med_IQR$coefficients[2])%>% 
  select(rowid, pred_IQR, opt_resid_IQR, calib_IQR)

lst_all_val <- lst_all_val %>% 
  full_join(., validation_IQR)

#plot to make sure calibration is correct
ggplot(lst_all_val, aes(x = temp_med, y = surface_temp_median)) +
  geom_point() +
  geom_point(aes(x = temp_med, y = calib_IQR), color = 'red')+
  geom_abline(intercept = val_med_IQR_dem$coefficients[1], slope = val_med_IQR_dem$coefficients[2], color = 'blue') +
  geom_abline(intercept = val_med_IQR_dem$ci[1,1], slope = val_med_IQR_dem$ci[2,1], color = 'blue', linetype = 2) +
  geom_abline(intercept = val_med_IQR_dem$ci[1,2], slope = val_med_IQR_dem$ci[2,2], color = 'blue', linetype = 2) +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(xlim = c(7,27),
                  ylim = c(7,27)) +
  labs(y = 'satellite-derived median\nskin temperature (deg C)', 
       x = expression(paste(italic('in-situ'), ' median water temp (deg C)'))) +
  final_theme

```

## look at the two predictions to see how they relate
```{r}
ggplot(lst_all_val, aes(x = calib_range, y = calib_IQR)) +
  geom_point() +
  geom_abline(intercept =  0, slope = 1) +
  final_theme

```

# Validation trials - play with in-situ database

## only use primary GLEON buoy data

## only use 2018 data

## only use littoral data (from KLC traps)

## only use 2 years of data (all sources)

## only use 2 years of data (buoy only)



# Validation trials - play with LS variables

Goal here: what is important to LS variables? How does it change the regression and error calcs?


## predicted v observed 50% cutoff for lake - range
```{r}
validtion_50_range <- validation_range %>% 
  filter(perc_lake > 50)
val_med_ls_50_range <- lm(validtion_50_range$surface_temp_median ~ validtion_50_range$temp_med)
summary(val_med_ls_50_range)

val_med_dem_50_range <-deming::deming(validtion_50_range$surface_temp_median ~ validtion_50_range$temp_med)
val_med_dem_50_range

cor(validtion_50_range$surface_temp_median, validtion_50_range$temp_med)

ggplot(validtion_50_range, aes(x = temp_med, y = surface_temp_median)) +
  geom_point() +
  geom_abline(intercept = val_med_dem_50_range$coefficients[1], slope = val_med_dem_50_range$coefficients[2], color = 'blue') +
  geom_abline(intercept = val_med_dem_50_range$ci[1,1], slope = val_med_dem_50_range$ci[2,1], color = 'blue', linetype = 2) +
  geom_abline(intercept = val_med_dem_50_range$ci[1,2], slope = val_med_dem_50_range$ci[2,2], color = 'blue', linetype = 2) +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(xlim = c(7,27),
                  ylim = c(7,27)) +
  labs(y = 'satellite-derived median\nskin temperature (deg C)', 
       x = expression(paste(italic('in-situ'), ' median water temp (deg C)'))) +
  final_theme
ggsave(paste0(figdir, 'deming_reg_median_50perc_range_v27Sept2021.jpg'), height = 5, width = 5, units = 'in')


```


## predicted v observed 50% cutoff for lake - IQR
```{r}
validtion_50_IQR <- validation_IQR %>% 
  filter(perc_lake > 50)
val_med_ls_50_IQR <- lm(validtion_50_IQR$surface_temp_median ~ validtion_50_IQR$temp_med)
summary(val_med_ls_50_IQR)

val_med_dem_50_IQR <-deming::deming(validtion_50_IQR$surface_temp_median ~ validtion_50_IQR$temp_med)
val_med_dem_50_IQR

cor(validtion_50_IQR$surface_temp_median, validtion_50_IQR$temp_med)

ggplot(validtion_50_IQR, aes(x = temp_med, y = surface_temp_median)) +
  geom_point() +
  geom_abline(intercept = val_med_dem_50_IQR$coefficients[1], slope = val_med_dem_50_IQR$coefficients[2], color = 'blue') +
  geom_abline(intercept = val_med_dem_50_IQR$ci[1,1], slope = val_med_dem_50_IQR$ci[2,1], color = 'blue', linetype = 2) +
  geom_abline(intercept = val_med_dem_50_IQR$ci[1,2], slope = val_med_dem_50_IQR$ci[2,2], color = 'blue', linetype = 2) +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(xlim = c(7,27),
                  ylim = c(7,27)) +
  labs(y = 'satellite-derived median\nskin temperature (deg C)', 
       x = expression(paste(italic('in-situ'), ' median water temp (deg C)'))) +
  final_theme
ggsave(paste0(figdir, 'deming_reg_median_50perc_IQR_v27Sept2021.jpg'), height = 5, width = 5, units = 'in')


```


## by cloud cover 50% - range

```{r}
validtion_50cc_range <- validation_range %>% 
  filter(cloud_cover < 50)
val_med_ls_50cc_range <- lm(validtion_50cc_range$surface_temp_median ~ validtion_50cc_range$temp_med)
summary(val_med_ls_50cc_range)

val_med_dem_50cc_range <-deming::deming(validtion_50cc_range$surface_temp_median ~ validtion_50cc_range$temp_med)
val_med_dem_50cc_range

cor(validtion_50cc_range$surface_temp_median, validtion_50cc_range$temp_med)

ggplot(validtion_50cc_range, aes(x = temp_med, y = surface_temp_median)) +
  geom_point() +
  geom_abline(intercept = val_med_dem_50cc_range$coefficients[1], slope = val_med_dem_50cc_range$coefficients[2], color = 'blue') +
  geom_abline(intercept = val_med_dem_50cc_range$ci[1,1], slope = val_med_dem_50cc_range$ci[2,1], color = 'blue', linetype = 2) +
  geom_abline(intercept = val_med_dem_50cc_range$ci[1,2], slope = val_med_dem_50cc_range$ci[2,2], color = 'blue', linetype = 2) +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(xlim = c(7,27),
                  ylim = c(7,27)) +
  labs(y = 'satellite-derived median\nskin temperature (deg C)', 
       x = expression(paste(italic('in-situ'), ' median water temp (deg C)'))) +
  final_theme
ggsave(paste0(figdir, 'deming_reg_median_50perccc_range_v27Sept2021.jpg'), height = 5, width = 5, units = 'in')

```

## by cloud cover 50% - IQR

```{r}
validtion_50cc_IQR <- validation_IQR %>% 
  filter(cloud_cover < 50)
val_med_ls_50cc_IQR <- lm(validtion_50cc_IQR$surface_temp_median ~ validtion_50cc_IQR$temp_med)
summary(val_med_ls_50cc_IQR)

val_med_dem_50cc_IQR <-deming::deming(validtion_50cc_IQR$surface_temp_median ~ validtion_50cc_IQR$temp_med)
val_med_dem_50cc_IQR

cor(validtion_50cc_IQR$surface_temp_median, validtion_50cc_IQR$temp_med)

ggplot(validtion_50cc_IQR, aes(x = temp_med, y = surface_temp_median)) +
  geom_point() +
  geom_abline(intercept = val_med_dem_50cc_IQR$coefficients[1], slope = val_med_dem_50cc_IQR$coefficients[2], color = 'blue') +
  geom_abline(intercept = val_med_dem_50cc_IQR$ci[1,1], slope = val_med_dem_50cc_IQR$ci[2,1], color = 'blue', linetype = 2) +
  geom_abline(intercept = val_med_dem_50cc_IQR$ci[1,2], slope = val_med_dem_50cc_IQR$ci[2,2], color = 'blue', linetype = 2) +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(xlim = c(7,27),
                  ylim = c(7,27)) +
  labs(y = 'satellite-derived median\nskin temperature (deg C)', 
       x = expression(paste(italic('in-situ'), ' median water temp (deg C)'))) +
  final_theme
ggsave(paste0(figdir, 'deming_reg_median_50perccc_IQR_v27Sept2021.jpg'), height = 5, width = 5, units = 'in')

```





# MSE and bias by month using opt_resid
```{r}

# calculate mse with optomized residuals
#all data
validation_range %>% 
  summarise(mse = round(sum((opt_resid^2))/n(), digits = 2),
            rmse = round(sqrt(mse), digits = 2),
            bias = round(sum(opt_resid)/n(), digits = 2),
            mae = round(sum(abs(opt_resid))/n(), digits = 2),
            n_obs = n()) %>% 
    write.csv(., file.path(datadir, 'LS_deming_prediction_stats_v20Jul2021.csv'))

#all data
validation_range %>% 
  group_by(LSmission) %>% 
  summarise(mse = round(sum((opt_resid^2))/n(), digits = 2),
            rmse = round(sqrt(mse), digits = 2),
            bias = round(sum(opt_resid)/n(), digits = 2),
            mae = round(sum(abs(opt_resid))/n(), digits = 2),
            n_obs = n()) %>% 
    write.csv(., file.path(datadir, 'LS_deming_prediction_bymission_stats_v20Jul2021.csv'))


# by month
validation_range %>% 
  mutate(month = as.numeric(format(date, '%m'))) %>% 
  group_by(month) %>% 
  summarise(mse = round(sum((opt_resid^2))/n(), digits = 2),
            rmse = round(sqrt(mse), digits = 2),
            bias = round(sum(opt_resid)/n(), digits = 2),
            mae = round(sum(abs(opt_resid))/n(), digits = 2),
            n_obs = n()) %>% 
  select(month, n_obs, mae, rmse, bias) %>% 
  write.csv(., file.path(datadir, 'LS_deming_prediction_stats_allmissions_v29June2021.csv'))

#by month and mission
validation_range %>% 
  mutate(month = as.numeric(format(date, '%m'))) %>% 
  group_by(month, LSmission) %>% 
  summarise(mse = round(sum((opt_resid^2))/n(), digits = 2),
            rmse = round(sqrt(mse), digits = 2),
            bias = round(sum(opt_resid)/n(), digits = 2),
            mae = round(sum(abs(opt_resid))/n(), digits = 2),
            n_obs = n())  %>% 
  select(month, LSmission, n_obs, mae, rmse, bias) %>% 
  pivot_longer(!c(month, LSmission), names_to = 'variable', values_to = 'values') %>% 
  mutate(variable = factor(variable, levels = c('n_obs', 'mae', 'rmse', 'bias'))) %>% 
  arrange(month, variable) %>% 
  pivot_wider(id_cols = c(LSmission,  values, variable),
              names_from = c(month, variable),
              values_from = values) %>% 
  arrange(LSmission) %>% 
  write.csv(., file.path(datadir, 'LS_deming_prediction_stats_v29June2021.csv'))
```


#### plot residuals by x

```{r}
lst_insitu_dem_forresid = mcreg(x = validation_range$temp_med, y = validation_range$surface_temp_median, method.reg = 'Deming')


ggplot(validation_range, aes(sample = opt_resid)) +
  geom_qq() +
  geom_qq_line()

ggplot(validation_range, aes (y = opt_resid, x = temp_med)) +
  geom_point() +
  labs(x = 'in-situ median water temp (deg C)',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme
ggsave(paste0(dir, 'figures/optresid_x_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = temp_med, color = month)) +
  geom_point(size = 2) +
  labs(x = 'in-situ median water temp (deg C)',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme +
  scale_color_colorblind()
ggsave(paste0(dir, 'figures/optresid_x_colorbymonth_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')
```

#### residuals by cloud cover
```{r}
ggplot(validation_range, aes (y = opt_resid, x = cloud_cover)) +
  geom_point() +
  labs(x = 'cloud cover (percent)',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme
ggsave(paste0(dir, 'figures/optresid_cloudcover_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = cloud_cover, color = month)) +
  geom_point(size = 2) +
  labs(x = 'cloud cover (percent)',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme +
  scale_color_colorblind()
ggsave(paste0(dir, 'figures/optresid_cloudcover_colorbymonth_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

```

#### residuals by pixel count
```{r}
ggplot(validation_range, aes (y = opt_resid, x = pixel_count)) +
  geom_point() +
  labs(x = 'pixel count',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme
ggsave(paste0(dir, 'figures/optresid_pixelcount_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = pixel_count, color = month)) +
  geom_point(size = 2) +
  labs(x = 'pixel count',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme +
  scale_color_colorblind()
ggsave(paste0(dir, 'figures/optresid_pixelcount_colorbymonth_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

```

#### residuals by mean sample depth
```{r}
ggplot(validation_range, aes (y = opt_resid, x = depth_avg)) +
  geom_point() +
  labs(x = 'mean in-situ depth (m)',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme
ggsave(paste0(dir, 'figures/optresid_insitudepth_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = depth_avg, color = month)) +
  geom_point(size = 2) +
  labs(x = 'mean in-situ depth (m)',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme +
  scale_color_colorblind()
ggsave(paste0(dir, 'figures/optresid_insitudepth_colorbymonth_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

```

#### residuals by month, year, doy
```{r}
ggplot(validation_range, aes (y = opt_resid, x = month)) +
  geom_point() +
  labs(x = 'month',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme
ggsave(paste0(dir, 'figures/optresid_month_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = month, color = pixel_count)) +
  geom_point(size = 2) +
  labs(x = 'month',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme 
ggsave(paste0(dir, 'figures/optresid_month_bypixelcount_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = month, color = cloud_cover)) +
  geom_point(size = 2) +
  labs(x = 'month',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme 
ggsave(paste0(dir, 'figures/optresid_month_bycloudcover_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = year)) +
  geom_point() +
  labs(x = 'year',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme
ggsave(paste0(dir, 'figures/optresid_year_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = year, color = pixel_count)) +
  geom_point(size = 2) +
  labs(x = 'year',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme 
ggsave(paste0(dir, 'figures/optresid_year_bypixelcount_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = year, color = cloud_cover)) +
  geom_point(size = 2) +
  labs(x = 'year',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme 
ggsave(paste0(dir, 'figures/optresid_year_bycloudcover_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = doy)) +
  geom_point() +
  labs(x = 'doy',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme
ggsave(paste0(dir, 'figures/optresid_doy_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = doy, color = pixel_count)) +
  geom_point(size = 2) +
  labs(x = 'doy',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme 
ggsave(paste0(dir, 'figures/optresid_doy_bypixelcount_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')

ggplot(validation_range, aes (y = opt_resid, x = doy, color = cloud_cover)) +
  geom_point(size = 2) +
  labs(x = 'doy',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme 
ggsave(paste0(dir, 'figures/optresid_doy_bycloudcover_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')


```

#### residuals by sun angle and azimuth
```{r}
ggplot(validation_range, aes (y = opt_resid, x = azimuth)) +
  geom_point() +
  labs(x = 'azimuth (degrees)',
       y = 'deming optimized residuals (deg C)') +
  geom_hline(yintercept = 0, color = 'black') +
  final_theme
ggsave(paste0(dir, 'figures/optresid_azimuth_v28Jun2021.jpg'), height = 5, width = 8, units = 'in')


```


#### correlation between doy / pixel count / cloud cover

Because there seems to be some functional dependency between these three, let's show the correlation of the 3 of them
```{r}
cor_pixel_doy = round(cor(x = validation_range$pixel_count, y = validation_range$doy), digits = 2)
pixel_doy <- ggplot(validation_range, aes(x = pixel_count, y = doy)) +
  geom_point() +
  annotate('text', label = paste0('r = ', cor_pixel_doy), y = 125, x = 8500) +
  labs(x = 'pixel count\ncontributing to mean LST', y='day of year') +
  final_theme+
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

cor_pixel_cloud = round(cor(x = validation_range$pixel_count, y = validation_range$cloud_cover), digits = 2)
pixel_cloud <- ggplot(validation_range, aes(x = pixel_count, y = cloud_cover)) +
  geom_point() +
  annotate('text', label = paste0('r = ', cor_pixel_cloud), y = 0, x = 8500) +
  labs(x = 'pixel count\ncontributing to mean LST', y='percent cloud cover\nin Landsat image') +
  final_theme+
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

cor_doy_cloud = round(cor(x = validation_range$doy, y = validation_range$cloud_cover), digits = 2)
doy_cloud <- ggplot(validation_range, aes(x = doy, y = cloud_cover)) +
  geom_point() +
  annotate('text', label = paste0('r = ', cor_doy_cloud), y = 0, x = 275) +
  labs(x = 'day of year\n', y='percent cloud cover\nin Landsat image') +
  final_theme+
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

plot_grid(pixel_doy, pixel_cloud, doy_cloud, labels = c('a', 'b', 'c'), nrow = 1)

ggsave(paste0(dir, 'figures/correlation_plot.png'), width = 12, height = 4)

```



